---
title: "P8105 HW3"
author: "Zeqi Li"
date: "2024-10-14"
output: github_document
---

```{r setup, include = FALSE}
library(p8105.datasets)
library(tidyverse)
```

# Problem 1 
## Load dataset
```{r load_data_p1}
data("ny_noaa")
head(ny_noaa)
```

The NY NOAA dataset contains `r nrow(ny_noaa)` rows and `r nrow(ny_noaa)` 
columns. This dataset describes the weather in New York using variables such 
as date of observation, precipitation, snowfall, snow depth, max temperature, 
and min temperature.

## Tidy up data
```{r tidy_data_p1}
ny_noaa =
  ny_noaa |> 
  mutate(tmax = as.numeric(tmax),
         tmin = as.numeric(tmin)) |> 
  separate(date, into = c("year", "month", "day"), convert = TRUE)
```
To tidy up the dataset, `mutate()` is first applied to convert the `tmax` and 
`tmin` from `chr` data type to `num` data type. Then, `separate()` is applied 
to split the `date` variable into `year`, `month`, and `day` columns.

```{r tidy_data_cont}
na = sum(is.na(ny_noaa))

snow_count = 
  ny_noaa |> 
    count(snow) |> 
    arrange(desc(n))
```
The most common values in the dataset is `NA`. There are `r na` 
`NA` values in this dataset. Another noticeable common value is the 0s in the 
`snow` column. There are `r pull(snow_count, n)[1]` 0s in that column, mainly 
because there are many days in a year when there is no snow in New York.


